### 2020-07-04 chapters/abstract.tex
Some Background feedback applied
mais cenas
Small corrections
Abstract

Data pattern shift detection has been researched within diverse areas and application domains and refers to the process of finding patterns in data that do not conform to expected or usual behavior. Accurate and timely detection of data pattern deviations allows for immediate measures to be taken, preventing loss of income and maintaining the companies' reputation among customers.
As such, many large companies take this into consideration and deploy data anomaly components in their systems. In this way, system administrators can proactively adjust and prevent potential service failures.

For example, Yahoo's anomaly detection system fulfills its purpose of monitoring and raising alerts on time-series events for several of Yahoo's use cases. At Microsoft, their custom, internally made anomaly detection system monitors millions of metrics coming from a variety of different internal services, enabling engineers to solve live on-site issues before they scale and become bigger ones.

Many real-time stream monitoring systems are static once deployed in a production environment. Engineers of said systems usually configure them under the assumption that future data flowing through the system follows roughly the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates.

With this thesis, we set out to design and evaluate a system that makes use of lightweight streaming analytical methods to detect deviations in data patterns while in the presence of large volumes of high velocity, highly skewed, seasonal data.

\vspace*{10mm}\noindent
\textbf{Keywords}: real-time, streaming, lightweight, monitoring, volume, velocity, variety, seasonal

>>>>

%The Abstract
%State the problem, your approach and solution, and the main contributions of the paper. Include little if any background and motivation. Be factual but comprehensive. The material in the abstract should not be repeated later word for word in the paper.

%Neste contexto, tínhamos este problema, desenvolvemos isto, avaliamos desta forma e chegamos a estas conclusões

% The problem
%Data pattern shift detection has been researched within diverse areas and application domains and refers to the process of finding patterns in data that do not conform to expected or usual behavior. Accurate and timely detection of data pattern deviations allows for immediate measures to be taken, preventing loss of income and maintaining the companies' reputation among customers. As such, many large companies take this into consideration and deploy data anomaly detection components in their systems. In this way, system administrators can proactively adjust and prevent potential service failures.

% The problem
Many real-time stream monitoring systems are static once deployed in a production environment. The engineers of those systems configure them under the assumption that future data flowing through the system roughly follows the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. Data pattern shift detection refers to the process of finding patterns in data that do not conform to expected or usual behavior. Accurate and timely detection of data pattern deviations allows for immediate measures to be taken. Thus, the problem at hand is to determine when to reconfigure the system, \textit{e.g.}, a Machine Learning model, based on an analysis of the drifts in the stream of data.

%Approach/Solution and %Main contributions
In this thesis, we design a method that makes use of lightweight streaming aggregations and distribution divergence functions to alert deviations in data patterns in real-time, while in the presence of large volumes of high velocity, highly skewed and seasonal data. We develop a two-phased and two-windowed method: first we perform a batch analysis on a reference window and then a stream analysis over a sliding streaming window. We aggregate the contents of both windows using an approximate histogram aggregation based on Exponential Moving Averages. We do this for each event and for each of its fields, also denominated features in our context. We measure the divergence between both reference and sliding window histograms for each feature with the Jensen–Shannon Divergence function, but other functions like the Wassertein distance could be used. For each feature, we find out the probability value of each divergence measure and apply the Holm-Bonferroni multiple test correction to find out which probability values are lowest and the associated features. We generate alerts based on a user-defined probability threshold, for each feature.

We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi feature analysis. Our method accurately detected the introduced anomalies in the synthetic dataset experiments while maintaining high throughput, for single and multi feature analysis. Experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we defer further investigation to future work and formulate a set of hypothesis that might explain it and hence are worthy of pursuing next.

Our set of experiments supports the claim that sliding window aggregations and distribution divergence methods can be combined to detect data pattern shifts in streaming scenarios with constant time and memory complexity.

\vspace*{10mm}\noindent
\textbf{Keywords}: streaming, real-time, lightweight, monitoring, volume, velocity, seasonal

### 2020-07-04 chapters/conclusions.tex
Abstract

\subsubsection*{Hypothesis Discussion}
In our synthetic data experiments, we were able to detect all of the introduced anomalies. Additionally, we were able to do it achieving very high streaming throughputs and low latencies. In the real data experiments, once again, we managed to achieve high troughputs and low latencies but the produced alerts were not very accurate. We defer further investigation to future work because of the time constraints of this thesis but leave open the possiblity that the approximate aggregation value is more inaccurate than what we thought it would be, due to the fact that the EMA tail holds all past events.

>>>>

\subsubsection*{Hypothesis Discussion}
In our synthetic data experiments, we were able to detect all of the introduced anomalies. Additionally, we were able to do it achieving very high streaming throughputs. In the real data experiments, once again, we managed to achieve high troughputs but the produced alerts were not very accurate. We defer further investigation to future work because of the time constraints of this thesis but leave open the possiblity that the approximate aggregation value is more inaccurate than what we thought it would be, due to the fact that the EMA tail holds all past events.
